\EVOSUITE at the SBFT 2023 Tool Competition

### 2023-03-15 papers/sbst23/paper_sbst23.tex
Pass over paper

\begin{abstract}
  \EVOSUITE is a search-based unit test generation tool for Java. This paper summarises the results and experiences of \EVOSUITE's participation at the 11th unit testing competition at SBFT 2023, where \EVOSUITE achieved the highest overall score.
\end{abstract}

@@

\EVOSUITE~\cite{FrA11c} is an automatic test generation tool for Java classes.
It takes the name of the target class as input, along with a Java classpath that
locates the class's compiled bytecode and that of its dependencies, and produces
a test suite in JUnit~4 format as output.

In a preprocessing phase, \EVOSUITE uses a basic static analysis to collect
information about the relevant classes, such as their constructors, methods
and fields. \EVOSUITE loads these classes into the Java Virtual Machine, and
applies various bytecode instrumentations while doing so. This enables \EVOSUITE
to produce test execution traces, and to avoid test flakiness by replacing
non-deterministic calls with deterministic, mocked versions.

@@

\EVOSUITE implements various evolutionary algorithms that operate on this
representation. Early versions of \EVOSUITE followed the \emph{Whole Test Suite} (WTS)
approach~\cite{GoA_TSE12}, where entire test suites are optimised with
respect to their overall code coverage. However, a subsequent study~\cite{ea_evaluation}
found the \emph{Dynamic Many-Objective Sorting Algorithm} (DynaMOSA)~\cite{dynamosa}
to yield superior results. Unlike WTS, DynaMOSA operates directly at the test case level,
and it is now the default algorithm in \EVOSUITE.

Irrespective of which algorithm is chosen, the search is always driven by fitness
functions. These encode coverage criteria, and \EVOSUITE can optimise for multiple
criteria at the same time. Core fitness functions are based on traditional heuristics
for code coverage, such as branch distance and approach level (see ~\cite{GoA_TSE12} for more details.)
Besides this, there are further fitness functions based on mutation
testing~\cite{emse14_mutation} and other basic criteria~\cite{rojas2015combining}.

A large share of the complexity of \EVOSUITE's code base is due to efforts to
avoid flakiness in the generated tests. On the one hand, this covers extensive
instrumentation and mocking. On the other hand, \EVOSUITE also compiles and
executes each test individually to avoid compilation errors (which may be the
result of bugs in \EVOSUITE) or flakiness caused by non-determinism in the
class under test.

@@

\EVOSUITE is currently undergoing a significant rewrite to improve the quality
of its internal code base, for example a better use of Java generics to aid
type safety. The latest version (1.2.0) contains the first fruits of this effort,
but also introduced regressions that still existed at the time the competition was
hold. For this reason, the same version (1.1.0) as in previous years was used.
The remaining setup was largely similar to the 2022 iteration: DynaMOSA~\cite{dynamosa}
was chosen as search algorithm, and \EVOSUITE was configured with the parameters
established in prior years, which mainly uses the tuned default
values~\cite{arcuri2013parameter} and default coverage criteria~\cite{rojas2015combining}.
%
% No new features were added in this version, but support
% for Java 9+ was added, several bugs were fixed, and a major refactoring was applied
% to improve the code quality with respect to the
% use of Java Generics internally. While several smaller bugfixes were applied in
% the run-up to the 2021 competition, no further optimisation or testing was
% applied for 2022.
%
However, in previous years the competition only rewarded code coverage, and this year
test case readability is also factored in. To account for this, configuration now
enables a test minimisation step, and filters regression assertions using mutation analysis.
%
The time budget was again distributed by allocating \SI{50}{\percent} of the overall time
set by the competition organisers for the search, and the other \SI{50}{\percent} equally to
the remaining phases (e.g., assertion generation, compile-check, flakiness-check).

@@

For the first time, the competition also conducted a human study with 13
participants to assess test case readability. For each of the projects \Collections,
\Threeten, and \Spatial, participants were presented a test case
generated by each of the six competing tools, covering the same method under
test. The participants were then asked to rank the tests per project
on a scale from 1 (best) to 5 (worst), in terms of readability. Overall,
\EVOSUITE achieved the 2nd best rating (rank 2.23 on average), only
preceded by \UtbotConcolic (2.13 on average). \EVOSUITE's rank distribution
is shown in Figure~\ref{fig:readability}.

Most participants found \EVOSUITE-generated tests to be readable. When
\EVOSUITE ranked best, they positively highlighted the thoroughness of
generated assertions, using edge cases as inputs, and the fact that tests
where short. When \EVOSUITE ranked worst, participants criticised the lack
of comments and descriptive test names, repetitiveness in the code, and
needless assertions. While \EVOSUITE's competition setup enabled test and
assertion minimisation, this is a computationally very expensive process,
and it is likely \EVOSUITE was sometimes unable to finish in time,
leaving tests unoptimised.

Among the tests for \Collections the longest one was produced by \EVOSUITE.
Interestingly, it ranked both highest and lowest, with one participant
stating it was ``the most complete'', while another suggested the code
being longer makes it ``more confusing''. But as a general trend, participants
seemed to prefer shorter, more cohesive tests with assertions that only check
one particular aspect of the method under test.

@@

This paper reports on the participation of the \EVOSUITE test generation tool
in the 11th SBST Java Unit Testing Tool Contest. On average, \EVOSUITE
achieved \avgConditionsCoverageRatioLong branch coverage,
\avgLinesCoverageRatioLong line coverage, and a mutation score of
\avgMutantsCoverageRatioLong, using a search budget of \budgetLong on the \cuts
classes considered for the competition. Overall, this results in a score of
\score, which is the highest score of all tools in the competition.

@@

%-------------------------------------------------------------------------
%\def\IEEEbibitemsep{5pt plus 1pt}
%\def\IEEEbibitemsep{6pt}
%\clearpage
\bibliographystyle{IEEEtranS}
\bibliography{papers}
\balance

>>>>

\begin{abstract}
  \EVOSUITE is an automated unit test generation tool for Java based
  on meta-heuristic search techniques. This paper summarises the
  results and experiences of \EVOSUITE's participation at the 11th
  unit testing competition at SBFT 2023, where \EVOSUITE achieved the
  highest overall score.
\end{abstract}

@@

\EVOSUITE~\cite{FrA11c} is an automatic test generation tool for Java classes.
It takes the name of the target class as input, along with a Java classpath that
locates the class's compiled bytecode and that of its dependencies, and produces
a test suite in JUnit~4 format optimised for code coverage as output.

In a preprocessing phase, \EVOSUITE uses a basic static analysis to collect
information about the relevant classes, such as their constructors, methods
and fields. \EVOSUITE loads these classes into the Java Virtual Machine, and
applies various types of  bytecode instrumentation while doing so. This enables \EVOSUITE
to produce test execution traces, and to avoid test flakiness by replacing
non-deterministic calls with deterministic, mocked versions.

@@

\EVOSUITE implements several different evolutionary algorithms that operate on this
representation. Early versions of \EVOSUITE followed the \emph{Whole Test Suite} (WTS)
approach~\cite{GoA_TSE12}, where entire test suites are optimised with
respect to their overall code coverage. A subsequent study~\cite{ea_evaluation}
introduced the \emph{Dynamic Many-Objective Sorting Algorithm} (DynaMOSA)~\cite{dynamosa} algorithm, which was shown
to yield superior results. Unlike WTS, DynaMOSA operates directly at the test case level,
and it is now the default algorithm in \EVOSUITE.

Irrespective of which algorithm is chosen, the search is always driven by fitness
functions. These encode coverage criteria, and \EVOSUITE can optimise for multiple
criteria at the same time. Core fitness functions are based on traditional heuristics
for code coverage, such as branch distance and approach level (see \EVOSUITE's original paper~\cite{GoA_TSE12} for more details.)
Besides this, there are further fitness functions based on mutation
testing~\cite{emse14_mutation} and other basic criteria~\cite{rojas2015combining}.

\EVOSUITE consists of a large code base, of which a large share is
dedicated to avoiding flakiness in the generated tests. This is
achieved using extensive instrumentation and mocking, as well as
post-processing; for example, \EVOSUITE compiles and executes each
test individually to avoid compilation errors (which may be the result
of bugs in \EVOSUITE) or flakiness caused by non-determinism in the
class under test.

@@

\EVOSUITE is currently undergoing a significant rewrite to improve the quality
of its internal code base, for example a better use of Java generics to aid
type safety. The latest version (1.2.0) contains the first fruits of this effort,
but also introduced regressions that still existed at the time the competition was
hold. For this reason, the same version (1.1.0) as in previous years was used.
The remaining setup was largely similar to the 2022 iteration: DynaMOSA~\cite{dynamosa}
was chosen as search algorithm, and \EVOSUITE was configured with the parameters
established in prior years, which mainly uses the tuned default
values~\cite{arcuri2013parameter} and default coverage criteria~\cite{rojas2015combining}.
%
% No new features were added in this version, but support
% for Java 9+ was added, several bugs were fixed, and a major refactoring was applied
% to improve the code quality with respect to the
% use of Java Generics internally. While several smaller bugfixes were applied in
% the run-up to the 2021 competition, no further optimisation or testing was
% applied for 2022.
%
While in previous years the competition only rewarded code coverage, this year
test case readability was also factored in. To account for this, the new configuration 
enables a test minimisation step, and filters regression assertions using mutation analysis.
%
The time budget was again distributed by allocating \SI{50}{\percent} of the overall time
set by the competition organisers for the search, and the other \SI{50}{\percent} equally to
the remaining phases (e.g., assertion generation, compile-check, flakiness-check).

@@

For the first time, the competition included a human study with 13
participants to assess test case readability. For each of the projects \Collections,
\Threeten, and \Spatial, participants were presented a test case
generated by each of the six competing tools, covering the same method under
test. The participants were then asked to rank the tests per project
on a scale from 1 (best) to 5 (worst), in terms of readability. Overall,
\EVOSUITE achieved the 2nd best rating (rank 2.23 on average), 
preceded by \UtbotConcolic (2.13 on average). \EVOSUITE's rank distribution
is shown in Figure~\ref{fig:readability}.

Most participants found \EVOSUITE-generated tests to be readable. When
\EVOSUITE ranked best, they positively highlighted the thoroughness of
generated assertions, using edge cases as inputs, and the fact that
tests where short. When \EVOSUITE ranked worst, participants
criticised the lack of comments and descriptive test names,
repetitiveness in the code, and needless assertions. While \EVOSUITE's
competition setup enabled test and assertion minimisation, this is a
computationally very expensive process, and it is likely \EVOSUITE was
sometimes unable to finish in time, leaving tests unoptimised. In
particilar, there are two aspects of this minimisation: The first
minimisation phase attempts to remove all statements in a test that
are irrelevant for achieving the test's coverage goals; the second
minimisation phase consists of selecting a minimal subset of test
assertions that are sufficient to detect all code mutants the test is
capable of detecting. If either of these phases exceeds the allocated
time budget it is cancelled, and \EVOSUITE returns the non-minimised
tests.

Among the tests for \Collections the longest one was produced by
\EVOSUITE.  Interestingly, it ranked both highest and lowest, with one
participant stating it was ``the most complete'', while another
suggested the code being longer makes it ``more confusing''. As a
general trend, participants seemed to prefer shorter, more cohesive
tests with assertions that only check one particular aspect of the
method under test.

@@

This paper describes the results and experiences of \EVOSUITE's
participation in the 11th SBST Java Unit Testing Tool Contest. On
average, \EVOSUITE achieved \avgConditionsCoverageRatioLong branch
coverage, \avgLinesCoverageRatioLong line coverage, and a mutation
score of \avgMutantsCoverageRatioLong, using a search budget of
\budgetLong on the \cuts classes considered for the
competition. Overall, this results in a score of \score, which is the
highest score of all competing tools.

@@

%-------------------------------------------------------------------------
%\def\IEEEbibitemsep{5pt plus 1pt}
%\def\IEEEbibitemsep{6pt}
%\clearpage
\balance
\bibliographystyle{IEEEtranS}
\bibliography{papers}
